---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"), stringsAsFactors = T)
data$pDate <- strptime(as.character(data$date), format = "%Y-%m-%d")
```



## What is mean total number of steps taken per day?
```{r}
total <- aggregate(data$steps, by = list(data$date), FUN = sum, na.rm = T)
names(total) <- c("Date", "TotalSteps")

hist(total$TotalSteps, main = "Total Steps per Day", xlab = "Steps")

means <- aggregate(data$steps, by = list(data$date), FUN = mean, na.rm = T)
names(means) <- c("Date", "Mean")
print(means)

medians <- aggregate(data$steps, by = list(data$date), FUN = median, na.rm = T)
names(medians) <- c("Date", "Median")
print(medians)

```


## What is the average daily activity pattern?


```{r}
intervals <- aggregate(data$steps, by = list(data$interval), FUN = mean, na.rm = T)
names(intervals) <- c("Interval", "Mean")
plot(intervals$Interval, intervals$Mean, type = "l", main = "Steps By Interval", xlab = "5 Minute Interval", ylab = "Mean")

```

Interval with Maximun number of Steps

```{r}
 print(intervals$Interval[intervals$Mean == max(intervals$Mean)])
```

## Imputing missing values

Total of missing Values

```{r}
sum(as.numeric(is.na(data$steps)))
```

Replacing NA with 5 minute interval mean

```{r}
new_data <- data;
indexes <- which(is.na(new_data))
new_intervals <-  new_data$interval[indexes]

for (i in 1:length(new_intervals)){
  new_data$steps[indexes[i]] <- intervals$Mean[intervals$Interval == new_intervals[i]]  
}


```

Recalculating Mean, Median and Plotting Histogram

```{r}
total <- aggregate(new_data$steps, by = list(new_data$date), FUN = sum)
names(total) <- c("Date", "TotalSteps")

hist(total$TotalSteps, main = "Total Steps per Day", xlab = "Steps")

means <- aggregate(new_data$steps, by = list(new_data$date), FUN = mean)
names(means) <- c("Date", "Mean")
print(means)

medians <- aggregate(new_data$steps, by = list(new_data$date), FUN = median)
names(medians) <- c("Date", "Median")
print(medians)
```


## Are there differences in activity patterns between weekdays and weekends?

Making names of days on English and Making a new Column for WeekDay type
```{r}
library(ggplot2)
Sys.setlocale("LC_TIME", "C")
data_weekdays <- weekdays(new_data$pDate);
weekend_indexes <- which(data_weekdays == "Saturday" | data_weekdays == "Sunday") 
weekday_indexes <-  which(data_weekdays != "Saturday" & data_weekdays != "Sunday")
new_data$weekDays <- data_weekdays
new_data$weekDays[weekend_indexes] <- "Weekend"
new_data$weekDays[weekday_indexes] <- "Weekday"
new_data$weekDays <- as.factor(new_data$weekDays)

g <- ggplot(new_data, aes(interval, steps))
p <- g + geom_line(aes(color = weekDays)) + 
  facet_grid(. ~ weekDays) + 
  labs(title = "Steps by Weekday", y = "Steps", x = "Interval") + 
  theme_bw(base_family = "Times") 
print(p)

```


